|

### 2021-12-03 07-Chp-Triangulation.Rmd
End of day:

```{r, echo = FALSE, warning=FALSE, message=FALSE}
source("R/helper.R")
knitr::read_chunk("R/05-Code-CPRD-Analysis.R")
knitr::read_chunk("R/07-Code-Triangulation-Analysis.R")
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # Info on knitting format
```

## Lay summary

@@

@@

## Methods
### Data sources {#tri-data-sources}

This triangulation exercise draws on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review of existing evidence presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). This evidence base is then supplemented by new evidence on the association of statin use with dementia outcomes in the CPRD (Chapter \@ref(cprd-analysis-heading)) and the association of lipids with dementia outcomes in previously unanalysed datasets accessed via the DPUK (Chapter \@ref(ipd-heading)).

Table \@ref(tab:thesisOverview-table) summarises the research each data source has attempted to address, the exposures and outcomes considered in each case, and the contribution of each chapter to the triangulation exercise presented here.

@@

@@

### Qualitative triangulation / narrative comparison {#qual-tri}

As discussed in the introduction to this Chapter (Section \@ref(triangulation-overview)), as part of a narrative synthesis of the evidence all information sources were initially grouped by outcome, and the findings from each source were narratively compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias assessments performed.

This analysis built on the detailed risk-of-bias assessments performed and presented as part of the systematic review (Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading)). Similarly, risk-of-bias assessments were performed for each new source of evidence presented in this thesis. The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias), but in summary, RCTS were assessed using the RoB 2 tool;[@sterne2019] NRSI using the ROBINS-I tool;[@sterne2016] NRSE using the ROBINS-E tool (see Section \@ref(rob-tools-nrse));[@morganr2020; @french2019] and MR studies using the Mamluk et al. tool.[@mamluk2020]

@@

### Quantitative triangulation

In addition to a qualitative discussion of the evidence, I attempted to integrate the numerical results of the multiple approaches under a proposed quantitative triangulation framework. This approach incorporates recent advancements in the way that bias in results is assessed (namely the move to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to illustrate how causal questions could be addressed under this framework.

@@

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardise effect directions
1. Assess and visualise the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Calculate bias-/indirectness-adjusted results and perform meta-analysis

@@

These questions were chosen for several reasons. Firstly, work presented in previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid faction on a specific outcome (e.g. Alzheimer's disease), rather than a composite outcome (e.g. all-cause dementia), focuses on an single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(hold)), and use of an composite all-cause outcome in the triangulation framework may mask this.

@@

@@

Risk of bias was assessed for each result as discussed in Section \@ref(qual-tri). For the sake of consistency and ease of computing, I mapped the different acceptable risk-of-bias judgements in each tool to a harmonised set: "Low", "Moderate", and "High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis.[@sterne2016]

@@

@@

(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null - bias "favouring the comparator" will always pull the effect estimate to the right. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases, defined as towards or away from the null.

@@

To aid with the quantitative triangulation exercise, I created a new method to visualise the extent, direction and type of bias in each result, enabling detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), these "bias direction" plots show the level (or extent) of bias across the domains for each result reported using coloured blocks. However, in contrast to the previous forest plots, symbols are used to illustrate the predicted absolute direction of bias in that domain. Distinct symbol pairs are used to denote additive and proportional biases.

@@

As an illustrative example, consider the first causal question of the interest in this exercise, the effect of LDL-c in mid-life on Alzheimer's disease. Here, studies of lipids levels in this time period would require minimal adjustment, whereas studies examining statin use (indirect exposure/intervention) in late-life (indirect population) would be down-weighted due to reduced relevance to the causal question.

@@

@@

<!-- IDEA Sensitivity analysis - set means to zero and only adjust for variance. Essentially adding uncertainty based on the amount of bias but not moving the effect estimate -->

@@

## Results

### Qualitative triangulation / narrative comparison

__[Question: the "strength of evidence" language here makes for fairly ambiguous reading - e.g. "very weak evidence". Should I also include the effect estimates from each approach, which will help to clarify, if it is a narrative synthesis?]__

I identified no consistent effect of any lipid fraction on any dementia outcome across the evidence sources presented in this thesis, through there was some suggestion of a protective effect of LDL-c lowering (via statin use) on all-cause dementia and Alzheimer's disease, and a harmful effect of raised triglyceride levels on vascular dementia.

Analysis of lipid levels across both the systematic review (Chapter \@ref(hold)) and the IPD analysis (Chapter \@ref(hold)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia. Similarly, both a meta-analysis of RCTs of statin use and of MR analyses of genetic instruments that mimic statin use found very weak evidence of an association, providing indirect evidence on the effect of LDL-c lowering on this outcome.

In contrast, the meta-analysis of NRSI of statin use found a slight protective effect, though many of the studies in this analysis were found to be at risk of immortal time bias, which would be expected to favour the intervention. Finally, analysis of the association of statin use with all-cause dementia in the CPRD (Chapter \@ref(hold)) found evidence of a harmful effect, though this is likely to be driven by the substantial confounding by indication identified in the vascular dementia subgroup.

For fibrates, an alternative lipid-regulating agent, there was some evidence of an increased risk of all-cause dementia in the CPRD, though this did not replicate in the meta-analysis of previous studies examining the association of fibrates and all-cause dementia.

Meta-analysis of lipid levels in the systematic review (Chapter \@ref(hold)) found extremely weak evidence for an association of any lipid fraction with Alzheimer's disease. Similarly, a meta-analysis of MR analyses of genetic instruments that mimic statin use found very weak evidence of an association.

In contrast, the meta-analysis of NRSI of statin use found a slight protective effect, though similar to the all-cause dementia outcome, many of the studies in this analysis were found to be at risk of immortal time bias. However, the CPRD analysis (Chapter \@ref(hold)) attempted to account for this potential bias through the use of time-varying treatment indicator. It found very weak evidence for an effect, though the potential for differential misclassification on the basis of the exposure limit the validity of this finding.

The IPD analysis (Chapter\@ref(hold)) provided no information on this outcome, as the sole cohort containing data on Alzheimer's disease (Whitehall II) had previously been analysed in respect to this outcome and so was included in the meta-analysis of NRSE.

There was a general absence of existing evidence on vascular dementia outcomes.  Meta-analysis of the small number of studies examing  lipid levels in the systematic review (Chapter \@ref(hold)) found extremely weak evidence for an association of any lipid fraction with vascular dementia. Similarly, a meta-analysis of statin use found very weak evidence for an effect on this outcome. The CPRD analysis (Chapter \@ref(hold)) found evidence for a harmful association of statin use with vascular dementia, though this finding is likely to be the result of severe confounding by indication, as identified through the use of control outcomes.

Finally, the IPD analysis (Chapter \@Ref(hold)) provided supporting evidence for the absence of an effect of any fraction on vascular dementia, with the sole exception of triglycerides. Raised triglycerides were associated with an increased incidence of vascular dementia across two previously unanalysed cohorts though there is the potential for uncontrolled confounding due to the limited information on covariates available. With respect to evidence on treatments for hypertriglyceridemia, examination of fibrates in the CPRD provided very weak evidence of an association with vascular dementia.

### Quantitative triangulation

@@

As an illustrative example, the bias and indirectness inherent to the estimate of the effect of statin use on Alzheimer's disease in the CPRD are considered here. For this result, there were three domains at greater than low risk of bias: Bias due to confounding, Bias due to definition of the outcome, and Bias in the selection of the reported result (Figure \@ref(fig:biasDirectionSingle)). In Domain 1 (bias due to confounding), the study did not adjust for _ApoE4_ status, which is predicted to mask any true association between statins and Alzheimer's disease (see Section \@ref(cprd-limitations) for a fuller discussion). As such, the bias is predicted to _favour the comparator_ and so the absolute direction of bias is to the right. In Domain 7 (bias in measurement of the outcome), there was a high risk of differential misclassification on the basis of the exposure. Here, history of statin use may make a diagnosis of vascular dementia more likely than Alzheimer's disease, and so for this outcome, this bias is predicted to _favour the experiment_ (in this case, an intervention). As such, the absolute direction of bias is to the left. In Domain 7 (bias in selection of the reported result), for consistency with the other NRSI, in the absence of a protocol the bias is assumed to be _away from the null_. Given the position of the effect estimate below the null, the absolute direction of bias in this domain is to the left.

@@

With respect to indirectness, the result was judged to be a low risk in the population and outcome domains. However, the fact the result refers to the effect of statins, which may have an effect on dementia outcomes through a mechanism other than LDL-c lowering, put it at high risk of indirectness in the intervention/exposure domain.

@@

The estimates obtained under the two scenarios of additive bias are plotted in Figure \@ref(fig:fpLdlAdComparison), and no substantial differences were observed between the them.

<!----------------------------------------------------------------------------->
(ref:fpLdlAdComparison-cap) Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease, using two different scenarios for the relationship of levels of bias. The left forest plot shows the result where the difference in mean value between the "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. In the right plot, the difference in mean value between "High" and "Moderate" is twice that between "Moderate" and "Low".

(ref:fpLdlAdComparison-scap) Comparison of bias-/indirectness-adjusted meta-analyses under two different scenarios of prior distributions of bias.

```{r fpLdlAdComparison, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpLdlAdComparison-cap)', out.width='100%', fig.scap='(ref:fpLdlAdComparison-scap)'}
knitr::include_graphics(file.path("figures/tri/fp_paired_midlife_ldl_ad_scenarios.png"))
```
<!----------------------------------------------------------------------------->

@@

Comparison of the unadjusted and bias-/indirectness-adjusted results for the effect of triglycerides on vascular dementia did not demonstrate a substantial difference (unadjusted `r unadj_effect_vad`; adjusted `r adj_effect_scenario1_vad`; Figure \@ref(fig:fpTgVad)), though again, heterogeneity between the adjusted results was greatly reduced (unadjusted $\tau^2$ = `r unadj_vad_tau2`, $I^2$ = `r unadj_vad_I2`; adjusted  $\tau^2$ = `r adj_vad_tau2`, $I^2$ = `r adj_vad_I2`). Similarly there was minimal difference between the results under the two scenarios of bias (Figure \@ref(fig:fpTgVaDComparison))

@@

<!----------------------------------------------------------------------------->
(ref:fpTgVaDComparison-cap) Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia, using two different scenarios for the relationship of levels of bias. The left forest plot shows the result where the difference in mean value between the "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. In the right plot, the difference in mean value between "High" and "Moderate" is twice that between "Moderate" and "Low".

(ref:fpTgVaDComparison-scap) Comparison of bias-/indirectness-adjusted meta-analyses under two different scenarios of prior distributions of bias.

```{r fpTgVaDComparison, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpTgVaDComparison-cap)', out.width='100%', fig.scap='(ref:fpTgVaDComparison-scap)'}
knitr::include_graphics(file.path("figures/tri/fp_paired_midlife_tg_vad_scenarios.png"))
```
<!----------------------------------------------------------------------------->

@@

In addition, it has sought to provide a generalised framework for quantitative triangulation, building on existing methods for systematic domain-based risk of bias assessment and bias-/indirectness-adjusted meta-analysis. To illustrate the method, I considered two causal questions: the effect of LDL-c at mid-life with Alzheimer's disease and the effect of triglycerides on vascular dementia. In the quantitative triangulation framework, there was very weak evidence for an effect in relation to either of the causal questions considered. The heterogeneity between results produced by different study designs contributing to the analysis was substantially reduced using this method, though this was largely due to a reduction in the precision of each individual result.

@@

@@

A key issue in this analysis is the accuracy of the assessment of bias and indirectness in each result. It has been widely demonstrated that the inter-rater agreement when using risk-of-bias tools is low with regards to the extnet of bias assinged to each domain[@hartling2011; @minozzi2019; @minozzi2020] Though not previously studied, given the difficulty in assessing the predicted direction of bias, agreement on this aspect of the tools is likely to be lower still.

A further issue in this regard is that rigourness of the tools may not be equivalent. This issue can be illustrated by the Ostergaard _et al._ study[@ostergaard2015] having low risk of bias across all domains and thus receiving a substantial proportion of the weight in the first case study (Figure \@ref(fig:fpLdlAd)). While the RoB suite of tools (RoB 2.0, ROBINS-I, ROBINS-E) were developed by an expert working group and have been widely used, the Mamluk _et al._ tool[@mamluk2020] used to assess risk of bias in Mendelian randomisation studies was designed by the authors for use in their own review. If the Mamluk tool is failing to adequately assess bias in Mendelian randomisation studies, then any analysis based on its assessments is likely to be subject to bias. This observation is particularly problematic given the default position of the ROBINS-I and ROBINS-E tool to require a "Moderate" judgement in the confounding domain. Under these conditions, it is possible that a well performed cohort study (low risk of bias across everything other that bias due to confounding) will be down-weighted versus a potentially poor MR study.

@@

A final limitation to this analysis is the synthesis of different effect estimates. As discussed in Section \@ref(hold), hazard ratios are not directly comparable to odds ratios, as they inherently account for person time-at-risk.[@mckenzie2019] As non-randomised studies (NRSI/NRSI) of dementia outcomes report hazard ratios and Mendelian randomisation studies report odds ratios, the resulting synthesis may be biased by the integration of these two different effect measures.

<!-- TODO Need to add some text here that says that assessment of the different domains  -->

__[Note: I would be particularly interested in your feedback here. It seems quantitative triangulation will frequently require integration of different effect measures if comparing e.g. NRSI reporting hazard ratios with MR reporting odds ratios. How should this be handled?]__

@@

@@

## References

>>>>

```{r, echo = FALSE, warning=FALSE, message=FALSE}
source("R/helper.R")
knitr::read_chunk("R/03-Code-Systematic-Review.R")
knitr::read_chunk("R/05-Code-CPRD-Analysis.R")
knitr::read_chunk("R/06-Code-IPD-Analysis.R")
knitr::read_chunk("R/07-Code-Triangulation-Analysis.R")
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # Info on knitting format
```

::: {.laybox data-latex=""}
## Lay summary {-}

@@

:::

@@

## Data sources {#tri-data-sources}

The triangulation exercises presented in this chapter draw on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review of existing evidence presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). This evidence base is then supplemented by new evidence on the association of statin use with dementia outcomes in the CPRD (Chapter \@ref(cprd-analysis-heading)) and the association of lipids with dementia outcomes in previously unanalysed datasets accessed via the DPUK (Chapter \@ref(ipd-heading)).

Table \@ref(tab:thesisOverview-table) summarises the research question each data source has attempted to address, the exposures and outcomes considered in each case, and the contribution of each chapter to the triangulation exercise presented here.

<!-- TODO Consider adding AD as an outcome to the IPD row, depending on which result you want to use. -->

@@

Additionally, both the qualitative and quantitative triangulation exercises used the the detailed risk-of-bias assessments performed and presented as part of the systematic review (Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading)). Similarly, risk-of-bias assessments were performed for each new source of evidence presented in this thesis. The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias), but in summary, RCTs were assessed using the RoB 2 tool;[@sterne2019] NRSI using the ROBINS-I tool;[@sterne2016] NRSE using an early adapted version of the ROBINS-E tool (see Section \@ref(rob-tools-nrse));[@morganr2020; @french2019] and MR studies using the Mamluk et al. tool.[@mamluk2020]

@@

## Qualitative (narrative) triangulation

```{r rctStatinDementia, include = FALSE, message=FALSE, warning=FALSE}
```

```{r Hypercholesterolemia, include = FALSE, message=FALSE, warning=FALSE}
```

```{r mrLipidsAD, include = FALSE, message=FALSE, warning=FALSE}
```

```{r LipidsSD, include = FALSE, message=FALSE, warning=FALSE}
```

```{r lipidFrationsSetup, include = FALSE, message=FALSE, warning=FALSE}
```

```{r obsStatins, include = FALSE, message=FALSE, warning=FALSE}
```

```{r obsFibrates, include = FALSE, message=FALSE, warning=FALSE}
```

```{r mrStatins, include = FALSE, message=FALSE, warning=FALSE}
```

```{r prepIPDFigures,include=F}
```

As part of a qualitative triangulation of the new and existing evidence identified by this thesis, all information sources were initially grouped by outcome, and the findings from each source were narratively compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias in each.

@@

Analysis of lipid levels across both the systematic review (Chapter \@ref(sy-rev-results-heading)) and the IPD analysis (Chapter \@ref(ipd-analysis-heading)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia (TC `r obsLipids$Dementia_TC$estimate`; LDL-c `r obsLipids$Dementia_LDL$estimate`; HDL-c `r obsLipids$Dementia_HDL$estimate`; triglycerides `r obsLipids$Dementia_TG$estimate`). Similarly, meta-analysis of studies identified by the review which examined a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$Dementia$estimate`). Finally, a meta-analysis of RCTs of statin use (`r rct_statin_acd`) and of MR analyses of genetic instruments that mimic statin use (`r mrStatin$Dementia$estimate`) both found weak evidence of an association, providing indirect evidence on the effect of lipid lowering on this outcome.

In contrast, the meta-analysis of NRSI of statin use found a slight protective effect (`r obsStatins$Dementia$estimate`), though many of the studies in this analysis were found to be at risk of immortal time bias, which would be expected to favour the intervention. Finally, analysis of the association of statin use with all-cause dementia in the CPRD (Chapter \@ref(cprd-analysis-heading)) found evidence of a harmful association (`r anydem_sta_text`), though this is likely to be driven by the substantial confounding by indication identified in the vascular dementia subgroup.

For fibrates, an alternative lipid-regulating agent, there was some evidence of an increased risk of all-cause dementia in the CPRD (`r anydem_fib_text`), though this did not replicate in the meta-analysis of previous studies examining the association of fibrates and all-cause dementia (`r obsFibrates$Dementia$estimate`).

Meta-analysis of lipid levels in the systematic review found extremely weak evidence for an association of any lipid fraction with Alzheimer's disease (TC `r obsLipids$AD_TC$estimate`; LDL-c `r obsLipids$AD_LDL$estimate`; HDL-c `r obsLipids$AD_HDL$estimate`; triglycerides `r obsLipids$AD_TG$estimate`). Similarly, meta-analysis of studies examining a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$AD$estimate`). A meta-analysis of MR analyses of genetic instruments that mimic statin use found weak evidence for an association of lipid lowering with Alzheimer's disease (`r mrStatin$AD$estimate`).

In contrast, the meta-analysis of NRSI of statin use found a slight protective effect on this outcome (`r obsStatins$AD$estimate`), though similar to the all-cause dementia outcome, many of the studies in this analysis were found to be at risk of immortal time bias. However, the CPRD analysis, which attempted to account for this potential bias through the use of time-varying treatment indicator, found weak evidence for an effect (`r probad_sta_text`). This validity of this finding was limited by the potential for differential misclassification on the basis of the exposure in the CPRD analysis.

Finally, the IPD analysis provided no additional information on the Alzheimer's disease outcome, as a previously published analysis examined the association of lipid levels with this outcome in the Whitehall II cohort. The published analysis was included in the meta-analysis of NRSE, the results of which are presented above.

There was a general absence of existing evidence on vascular dementia outcomes. Meta-analysis of published results was only possible for the total cholesterol lipid fraction (`r obsLipids$VaD_TC$estimate`). A meta-analysis of studies examining hypercholesterolemia also found weak evidence for an association with vascular dementia (`r obsHyperchol$VaD$estimate`). Similarly, a meta-analysis of statin use found weak evidence for an effect on this outcome (`r obsStatins$VaD$estimate`). The CPRD analysis found evidence for a harmful association of statin use (`r vasdem_sta_text`), though this finding is likely to be the result of severe confounding by indication related to vascular factors, as identified through the use of control outcomes.

The IPD analysis provided supporting evidence for the absence of an effect of any fraction on vascular dementia, with the sole exception of triglycerides. Raised triglycerides were associated with an increased incidence of vascular dementia across two previously unanalysed cohorts (`r ipd_vasdem`), though there is the potential for uncontrolled confounding due to the limited information on covariates available. With respect to evidence on treatments for hypertriglyceridemia, examination of fibrates in the CPRD provided weak evidence of an association with vascular dementia (`r vasdem_fib_text`).

## Quantitative triangulation

The qualitative comparison presented above identified no consistent effect of any lipid fraction on any dementia outcome across the evidence sources presented in this thesis. However, it is clear that this qualitative approach faces difficulties in interpretation when the number of individual results available is large, as discussed in the introduction to this chapter. This is true even when using summary estimates from meta-analyses of primary studies, which sacrifice valuable information on the biases inherent to each individual result.

In the following section, I propose and apply a novel quantitative triangulation framework to address these limitations. The absence of any clear signals across the evidence base does not make for particularly interesting case studies but I will nonetheless use the identified evidence to illustrate the novel methodological approach.

### Methods

In addition to the qualitative discussion of the evidence presented in Section (\@ref(qual-tri)), I attempted to integrate the numerical results of the multiple approaches under a proposed quantitative triangulation framework. This approach incorporates recent advancements in the way that bias in results is assessed (namely the move to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to illustrate how causal questions could be addressed under this framework.

@@

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardise effect directions
1. Specify idealised version of each study
1. Assess and visualise the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Calculate bias-/indirectness-adjusted results and perform meta-analysis

@@

These questions were chosen for several reasons. Firstly, work presented in previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid faction on a specific outcome (e.g. Alzheimer's disease), rather than a composite outcome (e.g. all-cause dementia), focuses on an single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(cprd-confounding-by-ind)), and use of an composite all-cause outcome in the triangulation framework may mask this.

@@

#### Specify idealised version of each study

Once the set of relevant results have been identified, an idealised study for each is described
An idealised study can be viewed as a replicate of the study producing the result of interest but with all sources of bias removed. The risk of bias is then assessed against this idealised study, while the causal question addressed by each idealised study is compared with the target causal question to define the indirectness of each result.

Note that all of the risk-of-bias tools used to assess non-randomised studies required specification of an idealised version of the study, against which bias was assessed. If using a risk-of-bias assessment tool that does not require this step, the idealised version of the study should be defined in advance of risk-of-bias assessment.

@@

The assessment of bias in each result is discussed in Section \@ref(tri-data-sources) above.For the sake of consistency and ease of computing, I mapped the different acceptable risk-of-bias judgements in each tool to a harmonised set: "Low", "Moderate", and "High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis.[@sterne2016]

@@

<!-- REVIEW in fact it is a mandatory part of new ROBINS-E tool, while only optional in RoB2 and ROBINS-I. Maybe reiterate that using a very early version of ROBINS-E -->

@@

(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null - bias "favouring the comparator" will always pull the effect estimate to the right, and can induce a change from a protective to harmful effect. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases, defined as towards or away from the null.

@@

To aid with the quantitative triangulation exercise, I created a new method to visualise the extent, direction and type of bias in each result, enabling detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), these "bias direction" plots show the level (or extent) of bias across the domains for each result reported using coloured blocks. However, in contrast to the previous forest plots, symbols are used to illustrate the predicted absolute direction of bias in that domain. Distinct symbol pairs are used to denote additive and proportional biases. Illustrations of these are provided in the Section \@ref(quant-tri-results).

@@

As an illustrative example, consider the first causal question of the interest in this exercise, the effect of LDL-c in mid-life on Alzheimer's disease. Here, studies of lipids levels in this time period would require minimal adjustment, whereas studies examining statin use (indirect exposure/intervention) in late-life (indirect population) would be adjusted and down-weighted due to reduced relevance to the causal question.

@@

<!-- TODO Need to be able to explain the order in which the biases are adjusted for -->

@@

<!-- TODO Should a fixed or random effects be used, because in theory you are adjusting for all potential causes of heterogeneity. -->

@@

### Results {#quant-tri-results}

@@

As an illustrative example of the process using a single result, the bias and indirectness inherent to the estimate of the effect of statin use on Alzheimer's disease in the CPRD (as analysed in Chapter \@ref(cprd-analysis-heading)) are considered here. For this result, there were three domains at greater than low risk of bias: Bias due to confounding, Bias due to definition of the outcome, and Bias in the selection of the reported result (Figure \@ref(fig:biasDirectionSingle)). In Domain 1 (bias due to confounding), the study did not adjust for _ApoE4_ status, which is predicted to mask any true association between statins and Alzheimer's disease (see Section \@ref(cprd-limitations) for a fuller discussion). As such, the bias is predicted to _favour the comparator_ and so the absolute direction of bias is to the right. In Domain 7 (bias in measurement of the outcome), there was a high risk of differential misclassification on the basis of the exposure. Here, history of statin use may make a diagnosis of vascular dementia more likely than Alzheimer's disease, and so for this outcome, this bias is predicted to _favour the experimental_ (in this case, an intervention). As such, the absolute direction of bias is to the left. In Domain 7 (bias in selection of the reported result), for consistency with the other NRSI, in the absence of a protocol the bias is assumed to be _away from the null_. Given the position of the effect estimate below the null, the absolute direction of bias in this domain is to the left.

@@

Indirectness was intentionally not displayed in the plot above, as these figures are created using `robvis` which is focused on creating visualisations of risk-of-bias data. In addition, I found it more useful to articulate the indirectness of a result in tabular format. With respect to this example result, I judged it to be a low risk in the population and outcome domains (Table \@ref(tab:singleIndirect-table)). However, the result refers to the effect of statins on Alzheimer's disease risk, rather than LDL-c levels directly, putting it at high risk of indirectness in the intervention/exposure domain. Assessing the predicted direction of indirectness was more challenging than assessing the predicted direction of bias. In this case, treatment with statins result in temporary lowering of lipid levels which and could thus underestimate the true effect of LDL-c lowering (indirectness towards the null). In contrast, statins may have an effect on Alzheimer's disease through a mechanism other than LDL-c lowering effect of statins and so the effect is overestimated (indirectness away from null). As such, I judged the direction of indirectness in this domain to be unpredictable.

<!-- TODO Consider timing of outcome as an issue - but is this a relevance or bias issue -->

&nbsp;<!----------------------------------------------------------------------->  
(ref:singleIndirect-caption) Assessment of indirectness for a single example study.

(ref:singleIndirect-scaption) singleIndirect

```{r singleIndirect-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

@@

No substantial differences were observed between the summary estimates obtained under the two scenarios of additive bias (Figure \@ref(fig:fpLdlAdComparison) in Appendix \@ref(appendix-tri-analysis)).

<!-- TODO Need to present overall summary estimate under two scenarios of bias and then send people to the Appendix. -->

@@

Comparison of the unadjusted and bias-/indirectness-adjusted results for the effect of triglycerides on vascular dementia did not demonstrate a substantial difference (unadjusted `r unadj_effect_vad`; adjusted `r adj_effect_scenario1_vad`; Figure \@ref(fig:fpTgVad)), though again, heterogeneity between the adjusted results was greatly reduced (unadjusted $\tau^2$ = `r unadj_vad_tau2`, $I^2$ = `r unadj_vad_I2`; adjusted  $\tau^2$ = `r adj_vad_tau2`, $I^2$ = `r adj_vad_I2`). Similarly there was minimal difference between the results under the two scenarios of bias (Figure \@ref(fig:fpTgVaDComparison) in Appendix \@ref(appendix-tri-analysis)).

<!-- TODO Need to present overall summary estimate under two scenarios of bias and then send people to the Appendix. -->

@@

@@

In addition, it has sought to provide a generalised framework for quantitative triangulation, building on existing methods for systematic domain-based risk of bias assessment and bias-/indirectness-adjusted meta-analysis. To illustrate the method, I considered two causal questions: the effect of LDL-c at mid-life with Alzheimer's disease and the effect of triglycerides on vascular dementia. In the quantitative triangulation framework, there was weak evidence for an effect in relation to either of the causal questions considered. The heterogeneity between results produced by different study designs contributing to the analysis was substantially reduced using this method, though this was largely due to a reduction in the precision of each individual result.

@@

<!-- TODO Also mention here that when the number of results is large, performing elictation for each separately in addition to the risk of bias assessments is going to be difficult. -->

@@

A key issue in this analysis is the accuracy of the assessment of bias and indirectness in each result. It has been widely demonstrated that the inter-rater agreement when using risk-of-bias tools is low with regards to the extent of bias assigned to each domain[@hartling2011; @minozzi2019; @minozzi2020] Though not previously studied, given the difficulty in assessing the predicted direction of bias, agreement on this aspect of the tools is likely to be lower still.

A further issue in this regard is that the rigour of the tools may not be equivalent. This issue can be illustrated by the Ostergaard _et al._ study[@ostergaard2015] having low risk of bias across all domains and thus receiving a substantial proportion of the weight in the first case study (Figure \@ref(fig:fpLdlAd)). While the RoB suite of tools (RoB 2.0, ROBINS-I, ROBINS-E) were developed by an expert working group and have been widely used, the Mamluk _et al._ tool[@mamluk2020] used to assess risk of bias in Mendelian randomisation studies was designed by the authors for use in their own review. If the Mamluk tool is failing to adequately assess bias in Mendelian randomisation studies, then any analysis based on its assessments is likely to be subject to bias. This observation is particularly problematic given the default position of the ROBINS-I and ROBINS-E tool to require a "Moderate" judgement in the confounding domain. Under these conditions, it is possible that a well-performed cohort study (low risk of bias across everything other that bias due to confounding) will be down-weighted versus a potentially poor MR study.

@@

A final limitation to this analysis is the synthesis of different effect estimates. As discussed in Section \@ref(sys-rev-analysis-overview), hazard ratios are not directly comparable to odds ratios, as they inherently account for person time-at-risk.[@mckenzie2019] As non-randomised studies (NRSI/NRSI) of dementia outcomes report hazard ratios and Mendelian randomisation studies report odds ratios, the resulting synthesis may be biased by the integration of these two different effect measures. This is primarily a concern for common outcomes, as when the outcome is rare, the odds, risk and hazard ratios approximate each other.

<!-- REVIEW Could add that ORs could be converted to RRs using a measure of "hyptothecial" risk -->

@@

<!-- REVIEW Good to mention direct adjustment for bias, e.g. Greenlands writings -->

@@
